The {\sc VSPEC} code: Spectroscopic phase curves of 3D exoplanet models in the presence of stellar variability

### 2024-06-03 src/tex/ms.tex
Copied through section 2.3 (faculae)

>>>>

\subsection{Stellar Spectra \label{subsec:spectra}}
Once the surface coverage is calculated, a composite spectrum is computed from a grid of PHOENIX stellar spectra
\citep{husser2013} As of \texttt{VSPEC 0.1}, we have spectra between 2300 K and 3900 K, with steps of
100 K. Each spectrum has $\log{g} = 5$ and solar metalicity.
The raw spectra span from 0.1 to 20 \micron~ with $5 \times 10^{-6}$ \micron~ steps, however binned
versions are pre-computed for faster runtimes.

\subsection{Spots \label{subsec:spots}}
Our star spot model is nearly entirely based on observations of the sun. Sun spots can be resolved and are well-studied,
whereas spots on other stars (e.g., M dwarfs) can only be observed indirectly. We therefore designed our spot model to mimic
sun spots but with parameterized values for spot temperature and lifetime that can be matched to observations of other stellar types and ages.

On the Sun, spots have two regions: the dark, central umbra and the lighter, surrounding penumbra. A detailed review of sunspot behavior,
including sizes and lifetimes, is \citet{solanki2003}, however the specific solar values are not relevant to this description. However,
the spot area as a function of time is
\begin{equation}
    A(t) = \left\{
    \begin{array}{lr}
        A_0 e^{(t-t_0)/\tau}, & \text{if } t \leq t_0 \\
        A_0 - W(t-t_0), & \text{if } t > t_0
    \end{array}
    \right\}
\end{equation}
where $A_0$ is the maximum area reached, $t_0$ is the time of the maximum, $\tau$ is the exponential growth rate, and $W$ is the linear decay rate.

{\sc VSPEC} also takes in parameters describing the population of spots, such as the mean of the (lognormal) peak area distribution
\citep{bogdan1988}, the equilibrium surface coverage, and the distribution style.

\subsection{Faculae \label{subsec:faculae}}
Faculae are magnetically-generated regions of the solar surface that usually appear as bright points near the limb; we employ the ``hot wall''
model \citep{spruit1976} where faculae are described as three-dimensional pores in the stellar surface with a hot, bright wall, and a
cool, dark floor, as shown in Figure \ref{fig:fac_struct}.

Their three-dimensional structure causes faculae to behave differently depending on their angle from disk-center. Close to the limb,
the hot wall is visible to the observer, and faculae appear as bright points. Near the center, however, the cool floor is exposed and
faculae appear dark. To consider this effect in the faculae lightcurve, we compute the fraction of the facula's normalized area -- the
area on the disk it would occupy as a flat spot -- that is occupied by each the hot wall and cool floor. This is done via numerical integral
along the radius of the spot.

\begin{equation}
    f_{wall} = \frac{\int _{-R}^{R} Z_{\rm eff} dr}{\int _{-R}^{R} Z_{\rm eff} dr + \int _{-R}^{R} R_{\rm eff} dr}
\end{equation}
where $Z_{\rm eff}$ is the apparent height of the wall
\begin{equation}
    Z_{\rm eff} = \left\{
    \begin{array}{lr}
        Z_w \sin{\alpha}, & \text{if } \alpha \leq \alpha_{\rm crit} \\
        2\sqrt{R^2 - r^2}\cos{\alpha}, & \text{if } \alpha > \alpha_{\rm crit}
    \end{array}
    \right\}
\end{equation}
and $R_{\rm eff}$ is the apparent width of the floor
\begin{equation}
    R_{\rm eff} = \left\{
    \begin{array}{lr}
        2\sqrt{R^2 - r^2} - Z_w\sin{\alpha}, & \text{if } \alpha \leq \alpha_{\rm crit} \\
        0, & \text{if } \alpha > \alpha_{\rm crit}
    \end{array}
    \right\}
\end{equation}
for facula radius $R$, depth $Z_w$, and angle from disk-center $\alpha$. $r$ in this numerical scheme
is defined as the distance from the center of the facula along the radial line connecting the facula center to the
disk center. $\alpha_{\rm crit}$ is the value of alpha at which the floor is no longer visible and is defined to be $\arctan{\frac{2\sqrt{R^2-r^2}}{Z_w}}$.

According to \citet{1997ApJ...484..479T}, faculae temperatures (of both the floor and wall) are dependent on the facula radius,
while depth appears to be constant. They also find that the smallest faculae have no visible floor, and that even at disk center they
appear as bright points. We parameterize the floor temperature to be
\begin{equation}
    \Delta T_{\rm eff, floor} = \left\{
    \begin{array}{lr}
    \text{Not visible}, & \text{if } r<r_{\rm min} \\
    m_{\rm floor}(r-r_{\rm min}) + \Delta T_{\rm eff, floor, 0}, & \text{if } r \ge r_{\rm min}
    \end{array}
    \right\}
\end{equation}

Where $r$ is the radius of the facula, $r_{\rm min}$ is the minimum radius where the floor is visible, $\Delta  T_{\rm eff, floor, 0}$ is
the difference between the floor and photosphere \teff at $r_{\rm min}$, and $m_{\rm floor}$ is the slope of the relationship with
units of [temperature] [length]$^{-1}$.

Similarly, the wall temperature is parameterized

\begin{equation}
    \Delta T_{\rm eff,wall} = m_{\rm wall}r + \Delta T_{\rm eff,wall,0}
\end{equation}

Where $\Delta T_{\rm eff,wall,0}$ is the temperature of a zero-radius facula and and $m_{\rm wall}$ is the slope of the radius-temperature
relationship with units of [temperature] [length]$^{-1}$.
These relationships can be defined by the user, for example setting $m=0$ for constant temperatures.

Facula lifetimes are defined as the time it takes to decay by $e^{-2}$. Because faculae grow and decay exponentially
at the same rate, each facula spends 1 lifetime with a radius greater than $e^{-1}$ of it's maximum. Each facula is born
and dies at $e^{-2}$ of its maximum, effectively existing in the code for two lifetimes.

\citet{hovis-afflerbach2022} suggest a typical facula lifetime to be on the order of 6 hours,
with a distribution that resembles a Poisson function. However, we choose a lognormal distribution
for both lifetime and maximum radius because it does not allow for these values to be 0. We choose to
correlate lifetime and radius so that they are determined by the normally distributed random variable $\mu$;
for each new facula, $\mu$ is randomly drawn and determines both the faclue lifetime and maximum radius.

\begin{figure}
    \centering
    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_model.png}}
    \gridline{\includegraphics[width=0.5\textwidth]{figures/facula_depth.pdf}}
    \script{facula_depth.py}
    \caption{
        ``Hot Wall'' model of faculae. Faculae structure causes their contrast to be dependent on their distance
        from the center of the disk. {\bf Top}: Depiction of a facula on the limb of a star. The hot wall is exposed
        to the observer causing the pore to appear bright. At disk center, the cool floor is most visible. {\bf Bottom}:
        The effects of depression depth and viewing angle on facula brightness. The 3D structure of faculae is most apparent
        when radius $\sim$ depth. A toy flux model was used to demonstrate the shape of these curves, but their magnitudes
        in practice depend on stellar spectral models.
        }
    \label{fig:fac_struct}
\end{figure}
